Synthesis of Rules for Network Intrusion Detectors

### 2019-10-22 paper/paper.tex
pass

\subsection{Step 1: Rule Maximization}
\pgfplotsset{width=5.5cm,compat=1.8}

The goal of the first step is to synthesize a rule with the maximum
number of options that satisfy the malicious traffic. A rule is
maximal if no other option can be added without making it miss the
traffic.

\begin{wrapfigure}[14]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}  
  \begin{tikzpicture}
    \begin{axis}[
        ybar stacked,
        enlargelimits=0.15,
        legend style={at={(0.5,-0.15)},
          anchor=north,legend columns=-1},
        ylabel={\#rules},
        symbolic x coords={gen, http, icmp, udp, tcp},
        xtick=data,
        nodes near coords,
        nodes near coords align={vertical},
      ]
      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
      \legend{relevant, irrelevant}
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:distribution-rules-protocol}Distribution of
    rules per protocol and relevance for triggering action.}
\end{wrapfigure}
It is important to note that options can be specific to a given
protocol or general and they can influence or not triggering an
action. Figure~\ref{fig:distribution-rules-protocol} shows the
distributions of these options per protocol and their relevance for
triggering rule actions. This plot indicates that the problem search
space is defined based on the protocol of the malicious traffic. For
instance, if \tname{} detects that the attack involves \CodeIn{http}
traffic, then 34 different rule options will be considered in the
creation of the rule.

\Fix{describe extraction...}

@@

Any rule created from a subset of rule options still captures the
traffic as every rule option we infer in this step is independent and
satisfied by the traffic.

>>>>

\subsection{Step 1: Rule Maximization}
\pgfplotsset{width=6cm,compat=1.8}
\pgfplotsset{every tick label/.append style={font=\tiny}}

The goal of the first step is to synthesize a rule with the maximum
number of options that satisfy the malicious traffic. A rule is
maximal if no other option can be added without missing the malicious
traffic.

\begin{wrapfigure}[16]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}
  \scalebox{1.05}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
%          ylabel={\#rules},
          symbolic x coords={gen, http, icmp, udp, tcp},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
        \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
        \legend{relevant, irrelevant}
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-rules-protocol}Distribution of
    number of rules per protocol and relevance for triggering action.}
\end{wrapfigure}
It is important to note that options can be specific to a given
protocol or general and they can influence or not triggering an
action. The purpose of rule options \CodeIn{msg}, for instance, is to
print on the output an informative message indicating what kind of
intrusion has been
observed. Figure~\ref{fig:distribution-rules-protocol} shows the
distributions of these options per protocol and their relevance for
triggering rule actions. We obtained this information
by\Fix{...explicar como...} \Fix{eh um pouco suspeito ter tantas regras
  irrelevantes...}This plot indicates that the problem search
space is defined based on the protocol of the malicious traffic. For
instance, if \tname{} detects by analyzing the format of network packets that the attack involves \CodeIn{http}
traffic, then 34 different rule options will be considered in the
creation of the rule.

To infer options, \tname{} parses the protocol message, looking for
fields in the message associated with options. For example, fields
\Fix{...and...}  can be used to determine the option
\Fix{...}. Unfortunately, inferring options at the field granularity
level is insufficient to capture \CodeIn{content} options (see
Section~\ref{sec:example-suricata-rules}), which refer to parts of the
payload of the message and cannot be directly extracted directly from
the fields. Handling this case is critically important as most rules
contain \CodeIn{content} options. Figure~\ref{fig:distribution-contents} shows
the histogram of number of these options per rule (up to 10) for the
Suricata public rule database \Fix{cite}. Rules without \CodeIn{content}
options is an exception and half of the rules contain at least two
\CodeIn{content} options.
\begin{wrapfigure}[14]{r}{0.56\textwidth}
  \centering
  \vspace{-3ex}
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,25) (1,10222) (2,4956) (3,5589)
          (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-contents}Histogram of number of
    \CodeIn{content} options per rule (up to 10).}
\end{wrapfigure}
To find the values associated with these options, \tname{} splits the
payload of the malicious message in tokens using natural language
delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
spaces. The rationale for this decision is that we observed, by
inspecting existing rules, that \CodeIn{content} options typically refer to
sequence of characters in the payload of messages separated by these
delimiters. \Fix{describe extraction of options not in the
  message. give examples! search only makes sense if this exists!}

@@

The rule reported by the first step satisfies the criterion to only
capture positive traffic, however it is ``too'' specific---it is
unable to capture different manifestations of the same attack as the
rule would unlikely match exactly the options we found for the input
message. Some options added to the rule during the first step are
irrelevant for determining the attack and should be removed.

The goal of this step is to create a \emph{minimal} rule that captures
\emph{only} the positive malicious traffic. A rule is minimal if
discarding any option results in capturing negative (\ie{}, benign)
traffic. Recall that the previous maximization step produces a rule
that captures the positive traffic and all options are satisifed by
the positive traffic. Consequently, any subset of these options also
captures the positive traffic. The minimization procedure we propose
in this step builds on that observation. It iteratively discards
options from the candidate rule until no other options can be
removed. The procedure uses negative traffic to guide the
search. Public databases of negative traffic with \Fix{thousands} of
messages exist \Fix{cite cite} and can be used to increase the
accuracy of the minimization.
